Understanding the role of different proxy selection methods in High-Dimensional Propensity Score Framework - A Plasmode simulation study

### 2024-09-12 Writeup/manuscript.Rmd
add the common rate table and corresponding description

\textcolor{blue}{\textbf{Summary results}: The dataset comprises 7585 individuals. Among these, the prevalence of the exposure is 0.488, while the prevalence of the outcome is 0.237.}

@@

```{r, echo=FALSE, fig.align="center", fig.height=6, fig.width=8, fig.cap="Figure presenting a comparison of Risk Differences (RD) and Odds Ratios (OR) with 95% confidence intervals for different methods used to evaluate the association between obesity and diabetes risk. The analysis is based on data from the National Health and Nutrition Examination Survey (NHANES) for the years 2013-2018. Methods are arranged by the number of variables used in the models.\\label{fig:comparison-plot}", message=FALSE, warning=FALSE}
library(ggplot2)
library(dplyr)

@@

\textcolor{brown}{I found the results in the table here is a little bit different from the value I got from the real data analysis. See the updated plot "comparison-plot-updated" and its code below:}

```{r, echo=FALSE, fig.align="center", fig.height=6, fig.width=8, fig.cap="Figure presenting a comparison of Risk Differences (RD) and Odds Ratios (OR) with 95% confidence intervals for different methods used to evaluate the association between obesity and diabetes risk. The analysis is based on data from the National Health and Nutrition Examination Survey (NHANES) for the years 2013-2018. Methods are arranged by the number of variables used in the models.\\label{fig:comparison-plot-updated}", message=FALSE, warning=FALSE}
library(ggplot2)
library(dplyr)

@@

@@

\begin{table}[htbp]
\centering
\caption{Computing Time for the Real-World Analysis for each Algorithm}
\label{tab:computing-time}
\begin{tabular}{lc}
\toprule
\textbf{Method} & \textbf{Computing Time}\\
\midrule
\textbf{Kitchen Sink} & 2.942 \\
\textbf{Bross formula} & 2.280 \\
\textbf{Hybrid (Bross and LASSO)} & 2.517 \\
\textbf{LASSO} & 2.540  \\
\textbf{Elastic Net} & 2.814 \\
\textbf{Random Forest (RF)} & 379.485 \\
\textbf{XGBoost} & 1.444 \\
\textbf{Forward selection (FS)} & 7.417 \\
\textbf{Backward elimination (BE)} & 12.794 \\
\textbf{Genetic algorithm (GA)} & 105.697 \\
\bottomrule
\end{tabular}
\end{table}

\textcolor{blue}{All methods, aside from RF and GA, exhibit relatively fast computing times. Although RF and GA are slower, their computational demands still remain within reasonable time limits.}

@@

- hdPS: High-dimensional Propensity Score
- NHANES: National Health and Nutrition Examination Survey
- OR: Odds Ratio
- RD: Risk Difference
- SE: Standard Error
- MSE: Mean Squared Error
- LASSO: Least Absolute Shrinkage and Selection Operator
- GA: Genetic Algorithm
- RF: Random Forest
- XGBoost: Extreme Gradient Boosting
- FS: Forward Selection
- BE: Backward Elimination
- CV: Cross-Validation
- RR: Relative Risk
- ElasticNet: A regularized regression method that combines LASSO and Ridge regression

>>>>

\textcolor{blue}{\textbf{Summary results}: The dataset comprises 7585 individuals. Among these, the prevalence of the exposure is 48.8\% or 0.488 , while the prevalence of the outcome is 23.7\% or 0.237.}

@@

```{r, echo=FALSE, fig.align="center", fig.height=6, fig.width=8, fig.cap="Figure presenting a comparison of Risk Differences (RD) and Odds Ratios (OR) with 95\\% confidence intervals for different methods used to evaluate the association between obesity and diabetes risk. The analysis is based on data from the National Health and Nutrition Examination Survey (NHANES) for the years 2013-2018. Methods are arranged by the number of variables used in the models.\\label{fig:comparison-plot}", message=FALSE, warning=FALSE}
library(ggplot2)
library(dplyr)

@@

\textcolor{brown}{I found the results in the table here is a little bit different from the value I got from the real data analysis. Also I found the cap of the figure not printing properly on my computer. I have to add double backslash before the pct symbol \%. If on other computers the double backslash causes some error maybe deleting it would help? See the updated plot "comparison-plot-updated" and its code below:}

```{r, echo=FALSE, fig.align="center", fig.height=6, fig.width=8, fig.cap="Figure presenting a comparison of Risk Differences (RD) and Odds Ratios (OR) with 95\\% confidence intervals for different methods used to evaluate the association between obesity and diabetes risk. The analysis is based on data from the National Health and Nutrition Examination Survey (NHANES) for the years 2013-2018. Methods are arranged by the number of variables used in the models.\\label{fig:comparison-plot-updated}", message=FALSE, warning=FALSE}
library(ggplot2)
library(dplyr)

@@

\textcolor{brown}{Updated the pairwise comparison table with adding the value for kitchen sink (ks). Though as the ks model contains all variables, adding the values for it would not add anything useful, I think it is weird to omit a method in the table without saying anything. Also here I changed the order/arrange of the methods, to match to the order/arrange of methods in the Methods section.}

\textcolor{blue}{Table \ref{tab:method-comparison-updated} presents a pairwise comparison of the number of proxy features shared between different variable selection methods used in the analysis. Each cell in the table indicates the count of common proxy variables selected by the method in the corresponding row and column. The diagonal cells, where the row and column methods are the same, represent the total number of proxy variables selected exclusively by each method. The first column displays the number of proxy variables shared between each method and the kitchen sink (KS) model. Given that the KS model includes all proxy variables, the values in the first column are identical to those in the diagonal cells for each row, which also presents the total number of proxy variables selected by the method in the corresponding row.}

\begin{table}[htbp]
\centering
\caption{Updated Comparison of variable overlap of selected proxies across different methods used to evaluate the association between obesity and diabetes}
\label{tab:method-comparison-updated}
\begin{tabular}{lcccccccccc}
\toprule
 & \textbf{KS} & \textbf{Bross} & \textbf{Hybrid} & \textbf{LASSO} & \textbf{EN} & \textbf{RF} & \textbf{XGB} & \textbf{FS} & \textbf{BE} & \textbf{GA} \\
\midrule
\textbf{Kitchen sink (KS)} & 142 & & & & & & & & & \\
\textbf{Bross formula} & 100 & 100 & & & & & & & & \\
\textbf{Hybrid (Bross and LASSO)} & 49 & 49 & 49 & & & & & & & \\
\textbf{LASSO} & 60 & 47 & 47 & 60 & & & & & & \\
\textbf{Elastic Net (EN)} & 69 & 54 & 48 & 60 & 69 & & & & & \\
\textbf{Random Forest (RF)} & 100 & 71 & 38 & 46 & 52 & 100 & & & & \\  
\textbf{XGBoost (XGB)} & 48 & 38 & 24 & 28 & 30 & 37 & 48 & & & \\
\textbf{Forward selection (FS)} & 59 & 45 & 41 & 51 & 54 & 45 & 25 & 59 & & \\
\textbf{Backward elimination (BE)} & 59 & 45 & 41 & 51 & 54 & 45 & 25 & 59 & 59 & \\
\textbf{Genetic algorithm (GA)} & 64 & 44 & 28 & 36 & 40 & 49 & 25 & 35 & 35 & 64 \\
\bottomrule
\end{tabular}
\end{table}

\textcolor{blue}{Table \ref{tab:proxy-common-pct-bross} presents the comparison of common proxy counts and rates between different methods with the bross formula hdPS. Column 1 indicates the number of proxy variables selected by each variable selection method. Column 2 shows the number of common features selected by each method and the Bross formula method. Column 3 demonstrates the percentage of the common features in the total number of features of each method.}

\textcolor{blue}{From Table \ref{tab:proxy-common-pct-bross}, it can be observed that, with the exception of the bross hdPS and random forest (RF) models, where the number of proxies was manually set to 100, and the kitchen sink (KS) model, which inherently includes all proxies, the number of proxy variables selected by all other models ranges from 49 to 69. The hybird method of bross and lasso hdPS, by definition, performs lasso proxy selection on the 100 proxy variables selected by bross, resulting in an inherent common rate of 1.00. For the remaining methods, the common rates are clustered around 0.74, with XGBoost exhibiting the highest common rate of 0.792 and GA presenting the lowest common rate of 0.688.}

\begin{table}[htbp]
\centering
\caption{Comparison of the count and percentage of proxy variables selected by each methods in common with that by the Bross formula hdPS}
\label{tab:proxy-common-pct-bross}
\begin{tabular}{lccc}
\toprule
\textbf{Method} & \textbf{Total Count} & \textbf{Common Count} & \textbf{Rate in Common} \\ 
\midrule
\textbf{Kitchen sink} & 142 & - & - \\
\textbf{Bross formula} & 100 & - & - \\
\textbf{Hybrid (Bross and LASSO)} & 49 & 49 & 1.000 \\
\textbf{LASSO} & 60 & 47 & 0.783 \\
\textbf{Elastic Net} & 69 & 54 & 0.783 \\
\textbf{Random Forest (RF)} & 100 & 71 & 0.710 \\  
\textbf{XGBoost} & 48 & 38 & 0.792 \\
\textbf{Forward selection (FS)} & 59 & 45 & 0.763 \\
\textbf{Backward elimination (BE)} & 59 & 45 & 0.763 \\
\textbf{Genetic algorithm (GA)} & 64 & 44 & 0.688 \\
\bottomrule
\end{tabular}
\end{table}

@@

\textcolor{blue}{Table \ref{tab:computing-time} presents the computing time for each method. All methods, aside from RF and GA, exhibit relatively fast computing times. Although RF and GA are slower, their computational demands still remain within reasonable time limits.}

\begin{table}[htbp]
\centering
\caption{Computing time for the real-world analysis for each algorithm}
\label{tab:computing-time}
\begin{tabular}{lc}
\toprule
\textbf{Method} & \textbf{Computing Time}\\
\midrule
\textbf{Kitchen Sink} & 2.942 \\
\textbf{Bross formula} & 2.280 \\
\textbf{Hybrid (Bross and LASSO)} & 2.517 \\
\textbf{LASSO} & 2.540  \\
\textbf{Elastic Net} & 2.814 \\
\textbf{Random Forest (RF)} & 379.485 \\
\textbf{XGBoost} & 1.444 \\
\textbf{Forward selection (FS)} & 7.417 \\
\textbf{Backward elimination (BE)} & 12.794 \\
\textbf{Genetic algorithm (GA)} & 105.697 \\
\bottomrule
\end{tabular}
\end{table}

@@

- hdPS: High-dimensional Propensity Score
- NHANES: National Health and Nutrition Examination Survey
- OR: Odds Ratio
- RD: Risk Difference
- SE: Standard Error
- MSE: Mean Squared Error
- \textcolor{blue}{KS: Kitchen Sink}
- LASSO: Least Absolute Shrinkage and Selection Operator
- \textcolor{blue}{ElasticNet / EN}: A regularized regression method that combines LASSO and Ridge regression
- RF: Random Forest
- XGBoost: Extreme Gradient Boosting
- FS: Forward Selection
- BE: Backward Elimination
- GA: Genetic Algorithm
- CV: Cross-Validation
- RR: Relative Risk
